Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-24 Paper/method.tex
..

The active subspace for the 19-dimensional problem was also computed using the regression-based
approach. As discussed earlier, a linear regression fit to the model output was used to
approximate the gradient using the sequence of steps outlined in Algorithm 1.2 
in~\cite{Constantine:2015}. 

@@

we provide a comparison of the square of individual components of the dominant eigenvector, obtained using the two 
strategies. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is also plotted in
Figure~\ref{fig:comp} (right) to illustrate convergence characteristics for the regression-based strategy.

@@

Individual components of the dominant eigenvector from the two strategies are observed to be in excellent

@@

to decrease with iterations, and required model evaluations at only 30 samples to converge using $\tau$~=~0.1.

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the igntion delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the two strategies is 
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

We extend the comparative assessment of the two approaches by estimating
the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
active subspace is 1-dimensional)
using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the gradient-based and gradient-free approaches are plotted in
Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
verification, we compare these estimates with those obtained using the
normalized DGSMs in~\eqref{eq:ndgsm}, as reported in~\cite{Vohra:2018}, in
Figure~\ref{fig:comp_as} (right). 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
for the 19 uncertain pre-exponents ($A_i$'s); $r$
denotes the number of eigenvectors in the active subspace.
Right: A plot of the screening metric ($\tilde{\nu}_i$) for individual $A_i$'s 
(adapted from~\cite{Vohra:2018}).}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores estimated using the two approaches agree favorably with each other as well as those
based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

The above comparisons indicate that the two strategies yield consistent results for the 19-dimensional
H$_2$/O$_2$ reaction kinetics problem. In other words, the gradient estimates using pertubation and
regression are reasonably close, and therefore, using the regression-based strategy being computationally
advantageous, would be preferred in this case. In the following section, we shift our focus to 
the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.

>>>>

The active subspace for the 19-dimensional problem was also computed using regression-based
estimates of the gradient that does not require model evaluations at neighboring points as
discussed earlier.

@@

we provide a comparison of the square of individual components of the dominant eigenvector, obtained using both, perturbation-based and regression-based estimates of the gradient of the ignition 
delay. Additionally, the quantity, $\max(\delta \hat{\mat{W}}_{1,j}^{(r)})$ is plotted in
Figure~\ref{fig:comp} (right) to illustrate convergence characteristics of Algorithm~\ref{alg:grad}.

@@

Individual components of the dominant eigenvector from the two approaches are observed to be in excellent

@@

to decrease with iterations, and requires model evaluations at only 30 samples to converge using $\tau$~=~0.1.

As mentioned earlier, the model output $f(\bm{\xi})$ i.e. the igntion delay in the
H$_2$/O$_2$ reaction in this case, varies
 predominantly in the active subspace. Hence, 
$f(\bm{\xi})$ can be approximated as $G(\hat{\mat{W}}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\hat{\mat{W}}_1^\top\bm{\xi}$, 
regarded as the \textit{sufficient summary plot} (SSP), obtained using the perturbation-based and
regression-based gradient estimates are
compared in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the 
perturbation-based and the regression-based strategies for computing the active subspace.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the response in ignition
delay based on the considered probability distributions for $A_i$'s is observed to be approximately linear in the
active subspace. This observation further indicates that the active subspace is 1-dimensional in this case.

We further estimate
the normalized activity scores for individual uncertain inputs ($\tilde{\nu}_{i,r}$; $r$=1 since the
active subspace is 1-dimensional)
using the components of the dominant eigenvector as shown in Algorithm~\ref{alg:grad}.
The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated
using the perturbation-based and regression-based approaches are plotted in
Figure~\ref{fig:comp_as} (left).  Additionally, for the purpose of
verification, we compare these estimates with those obtained using the
normalized DGSMs in~\eqref{eq:ndgsm}, as reported in~\cite{Vohra:2018}, in
Figure~\ref{fig:comp_as} (right). 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of normalized activity scores ($\tilde{\nu}_{i,r}$'s) 
for the 19 uncertain pre-exponents ($A_i$'s); $r$
denotes the number of eigenvectors in the active subspace.
Right: A plot of the screening metric ($\tilde{\nu}_i$) for individual $A_i$'s 
(adapted from~\cite{Vohra:2018}).}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores based on the two approaches for gradient estimation agree favorably 
with each other as well as those
based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

The above comparisons indicate that the gradient of the ignition delay with respect to the
uncertain $A_i$'s is reasonably approximated using both, perturbation-based and regression-based
approaches in this case. Since the number of model evaluations required in the case of
regression are much fewer, this approach seems favorable for the 19-dimensional problem. 
In the following section, we shift our focus to 
the higher-dimensional H$_2$/O$_2$ reaction kinetics application involving 33 uncertain inputs.
